\LARGE A double-sampling extension of the German National Forest Inventory for design-based small area regression estimation of timber volume resources on forest district levels \\ \vspace{2ex} \large Large scale application to the federal state of Rhineland-Palatinate \vspace{-1ex}

### 2018-02-08 Draft_Form/tab/modacc_modterms.tex
update 07_02_2018
update 07_02_2018
update_08_02_2018
update_08_02_2018

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Sun Jan 07 17:11:26 2018
\begin{table*}[!htbp]
	\begin{center}
	\caption{Accuracy metrics for submodels of final OLS regression model. Interaction terms are indicated by ':'. () give the respective values on the cluster level.} 
	\label{tab:modacc_modterms}
    {\small %
	\begin{tabular}{llccc}
  \hline
model terms & model & $R^2_{adj}$ & $RMSE_{cv}$ & $RMSE_{cv}\%$ \\ 
  \hline
meanheight + stddev + meanheight$^2$ +  & full model &  0.48 & 140.62  & 46.69 \\
treespecies + ALSyear + & & (0.53) &  (101.61) & (33.57) \\ 
meanheight:treespecies + \\ meanheight:ALSyear + meanheight:stddev + \\ stddev:ALSyear &&& \\ \\
meanheight + stddev + meanheight$^2$ + & reduced model  & 0.45  & 145.62 & 48.35 \\
ALSyear + meanheight:ALSyear + & & (0.49) & (106.08) & (35.40) \\
meanheight:stddev + stddev:ALSyear &&& \\ \\
\hline
\hline
\end{tabular}
}%
\end{center}
\end{table*}
%\endgroup

>>>>

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Sun Jan 07 17:11:26 2018
\begin{table*}[!htbp]
	\begin{center}
	\caption{Accuracy metrics for the two OLS regression models. Interaction terms are indicated by ':'. () give the respective values on the cluster level.} 
	\label{tab:modacc_modterms}
    {\small %
	\begin{tabular}{llccc}
  \hline
model terms & model & $R^2_{adj}$ & RMSE & RMSE\% \\ 
  \hline
meanheight + stddev + meanheight$^2$ +  & full model &  0.48 & 139.22  & 45.98 \\
treespecies + ALSyear + & & (0.53) &  (101.61) & (33.57) \\ 
meanheight:treespecies + \\ meanheight:ALSyear + meanheight:stddev + \\ stddev:ALSyear &&& \\ \\

meanheight + stddev + meanheight$^2$ + & reduced model  & 0.45  & 144.13 & 47.60 \\
ALSyear + meanheight:ALSyear + & & (0.49) & (106.08) & (35.40) \\
meanheight:stddev + stddev:ALSyear &&& \\ \\
\hline
\hline
\end{tabular}
}%
\end{center}
\end{table*}
%\endgroup

### 2018-02-08 Draft_Form/tex/Materials_and_Methods(2).tex
update_08_02_2018
update_08_02_2018

% ----------------------------- %
% PSMALL estimator:
\subsubsection{Pseudo Small Area Estimator (\psmall{})}
% \textbf{Pseudo Small Area Estimator (\psmall{})}\\

@@

The model selection constituted a major part of the current study, especially because several issues such as a) the heterogeneity in the remote sensing data, b) the identification of the optimal support sizes under angle count sampling and c) the incorporation of tree species information had to be addressed and investigated in order to identify the most accurate regression model realizable under the given data. We will here thus only provide a shortened summary of the extensive analysis carried out and refer to the study of \citet{hill2017a} if the reader is interested in more detailed information on the subjects.\par

The model with highest \adjrsq{} and lowest RMSE was achieved with the variables \meanheight{}, \meanheight{}$^2$, \stddev{} and \alsyear{} from the ALS data as well as the \treespecies{} variable from the classification map as main effect terms. Additionally, interaction terms between \meanheight{} and \alsyear{}, \stddev{} and \alsyear{} , \meanheight{} and \stddev{}, and \meanheight{} and \treespecies{} were included. The model yielded an \adjrsq{} of 0.48 and an RMSE of 140.62 \mha{} (46.7\%) on the plot level (table \ref{tab:modacc_modterms}, full model). While the study of \citet{hill2017a} only provide the model evaluations on the plot level, the two-phase estimators described in section \ref{sec:SAestimators} derive and apply the regression coefficients and the residuals on the aggregated cluster level. We thus re-evaluated the model as used in the estimators also on the cluster level and found the model accuracies to be higher than on the plot level (\adjrsq{} of 0.53 and RMSE of 101.61 \mha{} and 33.6\%). A substantial improvement of the model accuracy was achieved by the stratification to ALS acquisition years, which otherwise considerably lowered the \adjrsq{} due to increased unexplained variation in the data. An additional evaluation of the model residuals within each ALS acquisition year stratum (table \ref{tab:adj_r2_within}) revealed that a) the model accuracy substantially varied between the strata and b) high \adjrsq{} values of 0.59 even up to 0.66 on the cluster level were particularly achieved in acquisition year strata close to the date of the terrestrial survey (2010--2012). This effect was substantially diminished when dropping the \alsyear{} variable from the model.\par

@@

Discuss outliers etc. here ...

@@

%% NOTES:
%
% - we will here thus only provide a shortened summary of the regression model and refer to the study of \citet{hill2017a} if the reader is interested in more detailed information on the subjects.
%
% - 
% - differences: zero-plots included (basically, we here HAVE to use ALL plots for the regression model calibration --> because we use an internal model)
%   --> comment on the (non) removal of outliers / leverage plots: see review-answers --> plots stay in dataset with exception of truly erroneous local density variables
%       'plots kept in the sampling frame (i.e. they are used as x e s2 in the estimators) MUST also be kept in the modelling frame --> to insure the zero-mean-residual assumption'
%       (important: also leverage plots or "outliers" might have a good reason to be kept in the sampling / modelling frame --> think of sherman tree)
%
% for discussion: better than removing plots would be the the use of disturbance classes (Massey --> change estimation) in order to decrease the impact on plots where major cuttings took place
% between the ALS acquisition and the terrestrial survey. However, this is not possible if one height acquisition is available

>>>>

% ----------------------------- %
% PSMALL estimator:
\subsubsection{Pseudo Small Area Estimator (\psmall{})}
\label{sec:psmall}
% \textbf{Pseudo Small Area Estimator (\psmall{})}\\

@@

The model selection constituted a major part of the current study, especially because sophisticated challenges such as a) the heterogeneity in the remote sensing data, b) the identification of the optimal support sizes under angle count sampling and c) the incorporation of tree species information had to be addressed and investigated in order to identify the most accurate regression model realizable under the given data. We will here thus only provide a shortened summary of the extensive analysis carried out and refer to the study of \citet{hill2017a} if the reader is interested in more detailed information on the subjects.\par

The model with highest \adjrsq{} and lowest RMSE was achieved with the variables \meanheight{}, \meanheight{}$^2$, \stddev{} and \alsyear{} derived from the ALS data as well as the \treespecies{} variable derived from the classification map as main effect terms. Additionally, interaction terms between \meanheight{} and \alsyear{}, \stddev{} and \alsyear{} , \meanheight{} and \stddev{}, and \meanheight{} and \treespecies{} were included. The model yielded an \adjrsq{} of 0.48 and an RMSE of 140.62 \mha{} (46.7\%) on the plot level (table \ref{tab:modacc_modterms}, full model). While the study of \citet{hill2017a} only provide the model evaluations on the plot level, the two-phase estimators described in section \ref{sec:SAestimators} derive and apply the regression coefficients and the residuals on the aggregated cluster level. We thus re-evaluated the model as used in the estimators also on the cluster level and found the model accuracies to be higher than on the plot level (\adjrsq{} of 0.53 and RMSE of 101.61 \mha{} and 33.6\%). A stratification to the ALS acquisition years substantially improved the model accuracy and thus proved to be an effective means in accounting for the artificially introduced noise in the data set due to ALS quality variations and time-gaps between the ALS and the terrestrial survey. However, the stratification led to highly unbalanced datasets when stratifying according to the \treespecies{} variable. For this reason, a species individual modelling within each \alsyear{} stratum remained infeasible, but might have further improved the model accuracy. An additional evaluation of the model residuals within each ALS acquisition year stratum (table \ref{tab:adj_r2_within}) revealed that the model accuracies substantially varied between the strata (table \ref{tab:adj_r2_within}). \adjrsq{} values above the average  were particularly achieved in ALS acquisition years close to the terrestrial survey date (0.59 even up to 0.66 on the cluster level). This effect was substantially diminished when dropping the \alsyear{} variable from the model term.\par

@@

It should be mentioned that there is a major difference between the external and internal model approach concerning the treatment of outliers or leverage points for model fitting. In order to ensure the zero-mean-residual property of the \psmall{} and \extpsynth{} estimator under the internal model approach (section \ref{sec:psmall} and \ref{sec:extpsynth}), all observations of the terrestrial sample locations $x \in s_2$ used for estimation must also be included in the modelling frame. Excluding an observation from the model fit thus requires deleting it also from the sampling frame, which has to be regarded as an interference with the random sampling process the design-based inference relies on. For this reason, a removal of potential outliers or leverage points is, strictly speaking, only justified if the terrestrial response value or the explanatory variable value turns out to be truly erroneous (e.g. typos or erroneous measurements). If this is not the case, the removal of outliers or influential data points might increase the model fit, but to the cost of possible bias for the estimates (bias-variance trade-off).

We conducted a leverage and outlier detection on the plot level for the full regression model applying the usual criteria of \textit{Leverages} and \textit{Cook's Distance} as amongst others described in \citet{fahrmeir2013}. The Cook's Distance values $D_i$ did not exceed a value of 0.019, and thus were far apart from the often cited critical threshold of $D_i > 0.5$. However, the critical threshold of $2p/n$ for the leverage values (i.e. twice the average of the hat matrix' diagonal entries) was exceeded by 10\% (568) of the (5377) observations.

We conducted an Influence Analysis, i.e. leverage / outlier detection, for the 'final model'. We here considered the usually applied criteria of \textit{Leverages} and \textit{Cook's Distance} as amongst others described in Fahrmeir 2013 \{Regression: models, methods and applications. \textit{Springer Science \& Business Media}. page 164-167\}. The critical threshold of $2p/n$ (i.e. twice the average of the hat matrix' diagonal entries), was exceeded by 10\% of the observations. However, only 3\% of these leverage points were assigned to studentized residuals > 1 or < -1. Leaving these 3\% of points out and recalculating the final model lead to a $R^{2}_{adj}$ of 0.494 compared to 0.485 when including them in the regression. The Cook's Distance values $D_i$ did not exceed a value of 0.019, and thus were far apart from the often cited critical threshold of $D_i > 0.5$. Based on these findings, we decided not to remove observations from the modelling data set. We added a paragraph commenting on this issue of influential data points in section 3.3 'Final Regression Model', page 9 RHS line 690-707.

@@

%% NOTES:
%
% - differences: zero-plots included !!! (basically, we here HAVE to use ALL plots for the regression model calibration --> because we use an internal model)
%   --> comment on the (non) removal of outliers / leverage plots: see review-answers --> plots stay in dataset with exception of truly erroneous local density variables
%       'plots kept in the sampling frame (i.e. they are used as x e s2 in the estimators) MUST also be kept in the modelling frame --> to insure the zero-mean-residual assumption'
%       (important: also leverage plots or "outliers" might have a good reason to be kept in the sampling / modelling frame --> think of sherman tree)
%
% for discussion: better than removing plots would be the the use of disturbance classes (Massey --> change estimation) in order to decrease the impact on plots where major cuttings took place
% between the ALS acquisition and the terrestrial survey. However, this is not possible if one height acquisition is available
